---
title: "Exercise2"
author: "Kalyango Jovan"
date: "10/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages 
```{r, echo=FALSE, include=TRUE, warning=FALSE}
#install.packages("haven")
library(haven)
library(rjags)
library(mcmcplots)

```

# Loading Data

```{r}
Exercise_2_Data <- read_sav("Exercise 2 - Data.sav")
View(Exercise_2_Data)
```


# Initialization 

They are completely arbitrary values, so chosen that way, but i will test if i can use what the linear model gives in one of the chains.  
```{r}
model <- lm(attitude ~ agreeableness + extraversion, data = Exercise_2_Data)
summary(model)

```


```{r}
init1 <- list(b = c( 0.869, 0.259, -0.08), tau = 0.5)
init2 <- list(b = c(7, 10, -4), tau = 0.05)
init <- list(init1, init2)

```


Next, create a model object in R by means of the jags.model function. We need to specify
a model .txt-file, data and number of chains:
```{r}
model.def2 <- jags.model(file = "Exercise2Model1.txt", data = Exercise_2_Data, n.chains = 2)
```


```{r}
# Burn-in period 
update(object = model, n.iter = 1000)

# Sample from the posterior distribution
parameters <- c("b", "sigma")
samples <- coda.samples(model = model.def2,  variable.names = parameters, n.iter = 10000)
summary(samples)
```

## Convergence

```{r}
# History and density plot
plot(samples)

```


```{r}
# Autocorrelation plots
autocorr.plot(samples)
```

```{r}
# Gelman-Rubin diagnostic
gelman.plot(samples)

```

Only agreeableness $b[2]$ can be distinguished from zero based on the 95% CCI $[0.075, 0.44]$, so only agreeableness seems a reasonable predictor.

## G: Analysis with an interaction effect

Next, create a model object in R by means of the jags.model function. We need to specify
a model .txt-file, data and number of chains:
```{r}
model.defint <- jags.model(file = "Exercise2Modelinter.txt", data = Exercise_2_Data, n.chains = 2)
```


```{r}
# Burn-in period 
update(object = model, n.iter = 1000)

# Sample from the posterior distribution
parameters <- c("b", "sigma")
samples2 <- coda.samples(model = model.defint,  variable.names = parameters, n.iter = 10000)
summary(samples2)
```

$b[4]= 0.11$ cannot be distinguished from zero based on the 95% CCI $[-0.021, 0.24]$, so there is little evidence for an interaction effect.

```{r}
mcmcplot(mcmcout = samples)
```






